Discrete Quantum Theories and Computing

### 2018-11-26 dissertation.tex
dissertation.lyx "2.2 Quantum Probability" Updating...

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will actually happen. To maintain
the coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on the
both sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This can be achieved by mapping a classical event~$E$
to the projector generated by $E$
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$ because
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ are mapped to
the same projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Therefore, for any classical event~$E$, its corresponding quantum
event is defined to be the projector $\varphi\left(E\right)$, and
the set of all projectors on a given Hilbert space is called a quantum
event space~$\events$.

This function $\varphi$ not only respects the probability of events
but also naturally sends the set structure to the corresponding projector
structure: $\varphi\left(\Omega\right)=\mathbb{1}$ and $\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$.
Given two \emph{commuting} projectors $P_{0}$ and~$P_{1}$, there
exists a sample space $\Omega$ and a pair of events $E_{0}\subseteq\Omega$
and $E_{1}\subseteq\Omega$ such that $P_{0}=\varphi\left(E_{0}\right)$
and $P_{1}=\varphi\left(E_{1}\right)$. Conversely, given a pair of
events $E_{0}$ and $E_{1}$ in the same sample space~$\Omega$,
their corresponding quantum events $\varphi\left(E_{0}\right)$ and
$\varphi\left(E_{1}\right)$ are commuting and satisfying the following
properties: 
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,.\label{eq:pullback-properties}
\end{align}
Moreover, $E_{0}$ and $E_{1}$ are disjoint if and only if $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are orthogonal, where two projectors
$P_{0}$ and $P_{1}$ are called \emph{orthogonal} if $P_{0}P_{1}=\mathbb{0}$.

Then, a quantum probability space can be defined by a quantum event
space~$\events$ together with a quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
subject to the corresponding constraints \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012}:
\begin{align}
\mu\left(\mathbb{0}\right) & =0\,, & \mu\left(\mathbb{1}\right) & =1\,, & \mu\left(\mathbb{1}-P\right) & =1-\mu\left(P\right)\,,\label{eq:QuantumProbability}
\end{align}
and for each pair of orthogonal projectors $P_{0}$ and $P_{1}$:
\begin{equation}
\mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
Because $\varphi$ respects the probability of events, if we restrict
$\varphi$ on an event space $2^{\Omega}$, the function $\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$
defined by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$
is a classical probability measure and called the pullback of $\mu$
by $\varphi:2^{\Omega}\rightarrow\events$.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to a quantum probability measure in CQT. However, since the
variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, searching for a Born rule step-by-step here might provide
a idea what we should do in Sec.~\ref{sec:Toward-IVPM} and Chapter~\ref{chap:QIVPM}.
While the situations will become more delicate in the later sections,
we will fortunately find the unique Born rule in CQT here.

@@

\begin{thm}[Gleason's theorem]\label{cor:Gleason's}In a Hilbert
space $\Hilb$ of dimension $D\geq3$, given a quantum probability
measure $\mu:\events\rightarrow\left[0,1\right]$, there exists a
unique mixed state~$\rho$ such that $\mu\left(P\right)=\Tr\left(\rho P\right)$
for any $D$-dimensional projector~$P$.\end{thm}

Although Gleason's theorem can correspond every quantum probability
measure~$\mu$ to a unique mixed state~$\rho$, if we follow the
Stern-Gerlach type experiment interpretation in previous section,
we can find Gleason's theorem doesn't specify whether this state $\rho$
characterizing the state of particle sending to the quantum experiment.
For example, in contrast with CQT, a quantum theory could basically
ignore the input state and predict the experimental outcomes are equally
probable~\cite{doi:10.1142/S0217984913500644,1751-8121-47-40-405304}.
Even if a version of Bell's inequality could be proved in this quantum
theory, its prediction might be different from the prediction of CQT
enough that it is easily distinguish them experimentally.

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the corresponding quantum probability measure of the resulting
events. To define a sensible correspondence, we hope that if the state~$\ket{\Phi}$
is one of the outcomes of a quantum event~$P$, i.e., $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$
and vice versa. Moreover, since the physical phenomena exists and
should be the same no matter how we describe them, the probability
of an event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have $\muF_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muF_{\Phi}\left(P\right)$.
It is easy to check that the correspondence satisfying the conditions
in this paragraph is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. For mixed states $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$,
where $\ket{\Phi_{j}}\in\mathcal{H}$ are normalized, $q_{j}>0$,
and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule induces a
probability measure $\muB_{\rho}$ defined as 
\begin{equation}
\muB_{\rho}\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)\,.\label{eq:mixed-Born}
\end{equation}

>>>>

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will actually happen. To maintain
the coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on the
both sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This can be achieved by mapping a classical event~$E$
to the projector generated by $E$,
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$, because
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ are mapped to
the same projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Therefore, for any classical event~$E$, its corresponding quantum
event is defined to be the projector $\varphi\left(E\right)$, and
the set of all projectors on a given Hilbert space is called a quantum
event space~$\events$.

This function $\varphi$ not only respects the probability of events
but also naturally sends the set structure to the corresponding projector
structure: $\varphi\left(\Omega\right)=\mathbb{1}$ and $\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$.
Given two \emph{commuting} projectors $P_{0}$ and~$P_{1}$, there
exists a pair of events $E_{0}$ and $E_{1}$ in the same sample space~$\Omega$
such that $P_{0}=\varphi\left(E_{0}\right)$ and $P_{1}=\varphi\left(E_{1}\right)$.
Conversely, given a pair of events $E_{0}$ and $E_{1}$ in the same
sample space~$\Omega$, their corresponding quantum events $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are commuting and satisfying the
following properties: 
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,.\label{eq:pullback-properties}
\end{align}
Moreover, $E_{0}$ and $E_{1}$ are disjoint if and only if $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are orthogonal, where two projectors
$P_{0}$ and $P_{1}$ are called \emph{orthogonal} if $P_{0}P_{1}=\mathbb{0}$.

Then, a quantum probability space can be defined as a quantum event
space~$\events$ together with a quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
subject to the corresponding constraints \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012}:
\begin{align}
\mu\left(\mathbb{0}\right) & =0\,, & \mu\left(\mathbb{1}\right) & =1\,, & \mu\left(\mathbb{1}-P\right) & =1-\mu\left(P\right)\,,\label{eq:QuantumProbability}
\end{align}
and for each pair of orthogonal projectors $P_{0}$ and $P_{1}$:
\begin{equation}
\mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
Because $\varphi$ respects the probability of events, if we restrict
the domain of $\varphi$ on an classical event space $2^{\Omega}$,
the function $\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$
defined by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$
is a classical probability measure and called the pullback of $\mu$
by $\varphi:2^{\Omega}\rightarrow\events$.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to a quantum probability measure in CQT. However, since the
variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, searching for a Born rule step-by-step here might provide
a better idea what we should do in Sec.~\ref{sec:Toward-IVPM} and
Chapter~\ref{chap:QIVPM}. While the situations will become more
delicate in the later sections, we will fortunately find the unique
Born rule in CQT here.

@@

\begin{thm}[Gleason's theorem]\label{cor:Gleason's}In a Hilbert
space $\Hilb$ of dimension $D\geq3$, given a quantum probability
measure $\mu:\events\rightarrow\left[0,1\right]$, there exists a
unique mixed state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$ such
that $\mu\left(P\right)=\Tr\left(\rho P\right)$ for any $D$-dimensional
projector~$P$, where $\ket{\Phi_{j}}\in\mathcal{H}$ are normalized,
$q_{j}>0$, and $\sum_{j=1}^{N}q_{j}=1$.\end{thm}

To understand Gleason's theorem, we can follow the Stern-Gerlach type
experiment interpretation in the previous section. Although Gleason's
theorem correspond every quantum probability measure~$\mu$ to a
unique mixed state~$\rho$, it doesn't specify whether this state
$\rho$ characterizing the state of particle sending to the quantum
experiment. Consider an extreme example: a quantum theory could ignore
the input state and predict the experimental outcomes are always equally
probable. Even if these predictions form a quantum probability measure,
this kind of prediction is so different from the prediction of CQT
that they can be easily distinguished experimentally.

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the corresponding quantum probability measure of the resulting
events. To define a sensible correspondence, we hope that if the state~$\ket{\Phi}$
is one of the outcomes of a quantum event~$P$, $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$,
and vice versa. Moreover, since the physical phenomena exist and should
be the same no matter how we describe them, the probability of an
event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have $\muF_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muF_{\Phi}\left(P\right)$.
It is easy to check that the correspondence satisfying the conditions
in this paragraph is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. Since a mixed state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$
is a weighted average of projectors $\proj{\Phi_{j}}$ with weights
$q_{j}>0$ and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule
of $\rho$, $\muB_{\rho}\left(P\right)$, is also a weighted average
of $\muB_{\Phi_{j}}\left(P\right)$, 
\begin{equation}
\muB_{\rho}\left(P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)=\Tr\left(\rho P\right)\,,\label{eq:mixed-Born}
\end{equation}
which is also a quantum probability measure and consistent with Gleason's
theorem.
